Software reliability: Experiences\\ in European scientific research\\ projects and new trends

### 2017-11-08 paper.tex
cutting

The DataGrid \cite{cordis:datagrid} project (Jan 2001 -- Dec 2003) was devoted to develop software to provide Grid computing functionalities and related management tools for widely distributed scientific communities, such as the Large Hadron Collider (LHC) experiments, Earth observations and Bioinformatics. In this project, a large effort was dedicated in the definition and implementation of academic software engineering practices due to the unusual context in which DataGrid aims, starting from the large geographical distribution of the teams involved in the development tasks, evolving requirements coming from the different application areas and the design of a new technology: the Grid \cite{datagrid}. DataGrid developed its own software distribution, {\sl EDG} (European DataGrid), strongly based on Globus Middleware Services \cite{globus}.
%An Architecture Task Force coordinated the overall design and technical consistency of the developments, complemented by the definition of various testing phases: preparation, execution and documentation.

The three phases of Enabling Grids for E--sciencE (EGEE, Apr 2004 -- Apr 2010) \cite{cordis:egee, cordis:egee2, cordis:egee3} projects brought together scientists and engineers from more than 240 institutions in 45 countries worldwide to provide a seamless Grid infrastructure for e--Science. The {\sl gLite} middleware \cite{glite} was the ultimate official distribution of EGEE as of 2006, after two years of prototyping and re--engineering efforts.
%to converge with LHC Computing Grid ({\sl LCG-2}), Virtual Data Toolkit ({\sl VDT}) and {\sl Condor} \cite{condor} software distributions.
Starting in EGEE--II, there was a need to move to a more dynamic release process, and the first steps were taken towards an automated environment. At this stage no software metrics were collected, while functional and integration testing were performed manually, and only on the last project, EGEE--III, an automatic build system was adopted; the E--Infrastructure for Testing, Integration and Configuration of Software ({\sl ETICS}) service. The {\sl gLite} middleware developed within EGEE will evolve further as part of a new project: the European Middleware Initiative (EMI).

@@

The EMI (May 2010 -- Apr 2013) project \cite{cordis:emi} aimed to deliver a consolidated set of middleware products based on the four major middleware providers in Europe -- {\sl ARC}, {\sl dCache}, {\sl gLite} and {\sl UNICORE}. The EMI Quality Model used ISO/IEC 9126 Software Engineering -- Product Quality Standard \cite{iso9126} to identify a set of characteristics, each defining the associated metrics and Key Performance Indicators (KPIs). EMI was one of the first projects in using a Continuous Integration (CI) system for building, packaging and testing the various components. It leveraged the {\sl ETICS} tool and its plugin framework that provided not only the ability of collecting metrics during build and test execution, but also to make queries on the collected data and chart them through its chart generation framework.

The EGI--InSPIRE (Integrated Sustainable Pan--European Infrastructure for Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire} is the continuation of EGEE--III project, establishing and maintaining a sustainable European Grid Infrastructure. EGI--InSPIRE did not develop any software but maintained a production--ready Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution). The role of {\sl UMD} was to enforce the fulfilment of a set of quality criteria definitions \cite{egi-qc} in the software being delivered (i.e., {\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and deployed in the European scientific e--Infrastructures as a result of its continuation under a follow--up project, the EGI--Engage (Engaging the EGI Community towards an Open Science Commons) \cite{cordis:egi-engage}. This distribution is currently complemented by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.5\textwidth]{images/codestyle.png}
\caption{Code style standards followed by INDIGO--DataCloud's software products.}
\label{fig:fig_codestyle}
\end{figure}

The evaluation of the software quality is performed by measuring the values of the metrics and Key Performance Indicators (KPIs) defined based upon the ISO/IEC 9126 Software engineering -- Product quality standard. These metrics cover the development, release and maintenance phases of the software lifecycle.

@@

\bibitem{globus}
I. Foster, C. Kesselman, \emph{Globus: a Metacomputing Infrastructure Toolkit}, International Journal of Supercomputer Applications, vol. 11, no. 2, pp. 115–128, 1997

\bibitem{cordis:egee}
\emph{Enabling Grids for E-sciencE (EGEE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80149\_en.html

\bibitem{cordis:egee2}
\emph{Enabling Grids for E-sciencE-II (EGEE-II)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/99189\_en.html

\bibitem{cordis:egee3}
\emph{Enabling Grids for E-sciencE-III (EGEE-III)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/87264\_en.html

@@

\bibitem{cordis:etics}
\emph{E-Infrastructure for Testing, Integration and Configuration of Software (ETICS)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80138\_en.html

\bibitem{cordis:etics2}
\emph{E-Infrastructure for Testing, Integration and Configuration of Software - Phase 2 (ETICS 2)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/86604\_en.html

@@

\bibitem{cordis:emi}
\emph{European Middleware Initiative (EMI)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95311\_en.html

@@

\bibitem{cordis:egi-inspire}
\emph{Integrated Sustainable Pan-European Infrastructure for Researchers in Europe (EGI-InSPIRE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95923\_en.html

@@

\bibitem{cordis:egi-engage}
\emph{Engaging the EGI Community towards an Open Science Commons (EGI-ENGAGE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194937\_en.html

\bibitem{cordis:indigo}
\emph{INtegrating Distributed data Infrastructures for Global ExplOitation (INDIGO-DataCloud)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194882\_en.html

>>>>

The DataGrid \cite{cordis:datagrid} project (Jan 2001 -- Dec 2003) was devoted to develop software to provide Grid computing functionalities and related management tools for widely distributed scientific communities, such as the Large Hadron Collider (LHC) experiments, Earth observations and Bioinformatics. In this project, a large effort was dedicated in the definition and implementation of academic software engineering practices due to the unusual context in which DataGrid aims, starting from the large geographical distribution of the teams involved in the development tasks, evolving requirements coming from the different application areas and the design of a new technology: the Grid \cite{datagrid}.

%The three phases of Enabling Grids for E--sciencE (EGEE, Apr 2004 -- Apr 2010) \cite{cordis:egee, cordis:egee2, cordis:egee3} projects brought together scientists and engineers from more than 240 institutions in 45 countries worldwide to provide a seamless Grid infrastructure for e--Science. The {\sl gLite} middleware \cite{glite} was the ultimate official distribution of EGEE as of 2006, after two years of prototyping and re--engineering efforts. Starting in EGEE--II, there was a need to move to a more dynamic release process. At this stage no software metrics were collected, while functional and integration testing were performed manually, and only on the last project, EGEE--III, an automatic build system was adopted; the E--Infrastructure for Testing, Integration and Configuration of Software ({\sl ETICS}) service.i
The three phases of Enabling Grids for E--sciencE (EGEE, Apr 2004 -- Apr 2010) \cite{cordis:egee} projects brought together scientists and engineers from more than 240 institutions in 45 countries worldwide to provide a seamless Grid infrastructure for e--Science. The {\sl gLite} middleware \cite{glite} was the ultimate official distribution of EGEE as of 2006, after two years of prototyping and re--engineering efforts. Starting in EGEE--II, there was a need to move to a more dynamic release process. At this stage no software metrics were collected, while functional and integration testing were performed manually, and only on the last project, EGEE--III, an automatic build system was adopted; the E--Infrastructure for Testing, Integration and Configuration of Software ({\sl ETICS}) service.i

@@

The EMI (May 2010 -- Apr 2013) project \cite{cordis:emi} aimed to deliver a consolidated set of middleware products based on the four major middleware providers in Europe -- {\sl ARC}, {\sl dCache}, {\sl gLite} and {\sl UNICORE}. The EMI Quality Model used ISO/IEC 9126 Software Engineering -- Product Quality Standard \cite{iso9126} to identify a set of characteristics, each defining the associated metrics and Key Performance Indicators (KPIs). EMI was one of the first projects in using a Continuous Integration (CI) system, leveraging on {\sl ETICS} tool  for building, packaging and testing the various components.
%It leveraged the  and its plugin framework that provided not only the ability of collecting metrics during build and test execution, but also to make queries on the collected data and chart them through its chart generation framework.

The EGI--InSPIRE (Integrated Sustainable Pan--European Infrastructure for Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire} established a sustainable European Grid Infrastructure. EGI--InSPIRE maintained a production--ready Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution). The role of {\sl UMD} was to enforce the fulfilment of a set of quality criteria definitions \cite{egi-qc} in the software being delivered \cite{mario}. {\sl UMD} is still being used and deployed in the European scientific e--Infrastructures as a result of its continuation under a follow--up project, the EGI--Engage (Engaging the EGI Community towards an Open Science Commons) \cite{cordis:egi-engage}. This distribution is currently complemented by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

%\begin{figure}[!t]
%\centering
%\includegraphics[width=0.5\textwidth]{images/codestyle.png}
%\caption{Code style standards followed by INDIGO--DataCloud's software products.}
%\label{fig:fig_codestyle}
%\end{figure}

The evaluation of the software quality is performed by measuring the values of the metrics and KPIs defined based upon the ISO/IEC 9126 Software engineering -- Product quality standard. These metrics cover the development, release and maintenance phases of the software lifecycle.

@@

\bibitem{cordis:egee}
%\emph{Enabling Grids for E-sciencE (EGEE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80149\_en.html, EGEE-II -- http://cordis.europa.eu/project/rcn/99189\_en.html, EGEE-III -- http://cordis.europa.eu/project/rcn/87264\_en.html
\emph{EGEE} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80149\_en.html, EGEE-II -- http://cordis.europa.eu/project/rcn/99189\_en.html, EGEE-III -- http://cordis.europa.eu/project/rcn/87264\_en.html
%
%\bibitem{cordis:egee}
%\emph{Enabling Grids for E-sciencE (EGEE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80149\_en.html
%
%\bibitem{cordis:egee2}
%\emph{Enabling Grids for E-sciencE-II (EGEE-II)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/99189\_en.html
%
%\bibitem{cordis:egee3}
%\emph{Enabling Grids for E-sciencE-III (EGEE-III)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/87264\_en.html

@@

\bibitem{cordis:etics}
\emph{ETICS} project, CORDIS Europe, ETICS -- http://cordis.europa.eu/project/rcn/80138\_en.html, ETICS-2 -- http://cordis.europa.eu/project/rcn/86604\_en.html
%\bibitem{cordis:etics}
%\emph{E-Infrastructure for Testing, Integration and Configuration of Software (ETICS)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/80138\_en.html
%
%\bibitem{cordis:etics2}
%\emph{E-Infrastructure for Testing, Integration and Configuration of Software - Phase 2 (ETICS 2)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/86604\_en.html

@@

\bibitem{cordis:emi}
%\emph{European Middleware Initiative (EMI)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95311\_en.html
\emph{EMI} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95311\_en.html

@@

\bibitem{cordis:egi-inspire}
%\emph{Integrated Sustainable Pan-European Infrastructure for Researchers in Europe (EGI-InSPIRE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95923\_en.html
\emph{EGI-InSPIRE} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/95923\_en.html

@@

\bibitem{cordis:egi-engage}
%\emph{Engaging the EGI Community towards an Open Science Commons (EGI-ENGAGE)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194937\_en.html
\emph{EGI-ENGAGE} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194937\_en.html

\bibitem{cordis:indigo}
%\emph{INtegrating Distributed data Infrastructures for Global ExplOitation (INDIGO-DataCloud)} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194882\_en.html
\emph{INDIGO-DataCloud} project, CORDIS Europe, http://cordis.europa.eu/project/rcn/194882\_en.html
